Modeling ice-shelf cavities using Lagrangian elements

### 2017-06-13 Lagrangian_MISOMIP.tex
Working on JAMES revisions and Lagrangian paper

The calving of large tabular icebergs away from Antarctic ice shelves is an important feature of the Antarctic environment, and accounts for approximately half of Antarctic ice-shelf decay. 
However, despite its importance, the current generation of ice-ocean models is unable to represent ice-shelf calving in a physically realistic way.
One reason for this is that current ice-shelf models are mostly built on static Eularian grids, and were not designed to easily allow pieces of the ice shelf to break off and become icebergs. 
In this study we develop a new ice-shelf cavity model where the ice shelf is constructed out of Lagrangian elements that are bonded together by numerical bonds. This Lagrangian framework allows for large pieces of the ice shelf to break away and become tabular icebergs. 
We test newly developed Lagrangian ice -shelf model by simulating the circulation within a (static) idealized ice-shelf cavity, which was developed as part of the Marine Ice Ocean Modeling Inter-comparison Project (MISOMIP). 
The Lagrangian model results are compared to a second simulation that uses an Eularian ice shelf with an otherwise identical configuration.
Results show that the Lagrangian and Eularian models are almost indistinguishable.
% For example, differences in the models are two orders of magnitude smaller than the differences observed when changing the ocean model coordinate system. 
The similarly between the Eularian and Lagrangian models in a static ice-shelf configuration provides a proof of concept for the Lagrangian model, and means that we can confidently use Lagrangian ice-shelf models to extend the capabilities of the Eularian models.% The Lagrangian model is then used to simulate a tabular iceberg calving away from the ice shelf and drifting into the open ocean.

@@

%  and tables.
\begin{article}
%% ------------------------------------------------------------------------ %%
%
%  TEXT
%
%% ------------------------------------------------------------------------ %%
\section{Introduction}
Floating ice shelves cover vast regions of the Antarctic polar oceans. These massive platforms of ice extend deep into the water column, applying large pressures to surface of the ocean, which is often hundreds of meters below global sea level. Beneath the ice shelves, both the bottom topography and the ice-shelf geometry play a role in steering ocean currents \citep{Nicholls1996, Jenkins2010,  Stern2014}. The topographic constraint imposed by the ice shelves at the ocean's upper boundary significantly affects the circulation within the ice-shelf cavities, and gives the ocean in the region a unique character.

Satellite observations suggest that ice-shelf decay occurs via two main processes: melting and breaking \citep{Depoorter2013, Rignot2013}. Each of these is responsible for approximately half of the ice-shelf decay, and each influences the surrounding ocean (and ice-shelf geometry) in a distinct way.
Melting at the base of ice shelves causes freshwater fluxes into the ice-shelf cavity. The input of buoyant meltwater creates rising density plumes, which are guided along the ice-shelf base, and help drive ocean circulation beneath the ice shelves \citep{MacAyeal1984,  Holland2006}. Over time, melting at the ice-shelf base can erode the ice shelf, gradually altering the ice-shelf geometry.
%The strength of the circulation within the cavity feeds back onto the ice-shelf melt rates, by removing cold water from the cavity, and drawing in warmer waters from the open ocean, thus providing the constant supply of thermal energy needed for continuous ice-shelf melt \citep{Lewis1986, Jacobs2011}. 
In contrast, ice-shelf breaking causes sudden changes to the ice-shelf geometry, and releases giant icebergs into the ocean. After calving, these tabular icebergs can travel large distances and impact ocean hydrography \citep{Martin2010, Stern2015}, sea-ice formation \citep{Robinson2012, Stern2016} and ocean biology \citep{Smith2007, Vernet2012, Biddle2015} many miles away.

In recent years society's need for improved projections of future sea level has lead to an increased focus on the Antarctic mass balance, and hence, on modeling Antarctic sub-ice-shelf cavities \citep{Asay-Davis2016}.
% This increased interest has led to accelerated efforts to accurately model Antarctic ice-shelf cavities \citep{Asay-Davis2016}.
Modeling the ocean beneath the ice shelves presents a unique set of challenges, since (i) the presence of ice shelves provides a rigid upper boundary for the ocean model which is not encountered elsewhere in the ocean, and (ii) melting and breaking ice shelves imply a changing ocean boundary conditions which present numerous numerical difficulties.

The earliest models of ocean ice-shelf cavities were developed using static ice shelves with a fixed shape \citep{Hellmer1989, Determan1994, Grosfeld1997, Holland2001, Losch2008}. In these models, ice-shelf melting was represented through salinity and temperature fluxes, while the ice-shelf geometry remained unchanged. Later models of ice-shelf cavities allowed the ice-shelf geometry to evolve as the ice shelf melted, permitting the study of coupled ocean-ice phenomena \citep{Gladish2012, Sergienko2013}. More recently, dynamic ice-shelf models have been coupled the ocean cavity, allowing the study of grounding line migration, which is of key importance for sea level rise projections \citep{Grosfeld2004, Goldberg2012, De_Rydt2006}.

As far as we know, all models of ice-shelf cavities to date have omitted ice-shelf breaking and iceberg detachment. This is because (i) there is much uncertainty of the physics that govern ice-shelf breaking \citep{Benn2007, Alley2008, Levermann2012, Bassis2013}, and (ii) current models of ice-shelf cavities represent the ice shelves on static Eularian grids, which do not lend themselves to modeling iceberg detachment and drift. In contrast, existing {\it iceberg} models represent icebergs as Lagrangian particles, since this is a convenient way to model discrete objects traveling over large distances \citep{Bigg1997, Gladstone2001, Martin2010, Marsh2015}. To date there has been no real effort to synthesize these two approaches (i.e.: to combine ice shelf and iceberg models).

In this study we develop a new ice-shelf cavity model where the ice shelf is represented on a `Lagrangian grid'. In this model, the ice shelf is constructed out of Lagrangian elements which are bonded together by numerical bonds (see schematic Figures \ref{fig:Schematic}). This Lagrangian framework allows for large pieces of the ice shelf to break away and become tabular icebergs. An example of this enhanced capability of the Lagrangian model is demonstrated in Figures \ref{fig:Iceberg_calving}, which shows a tabular iceberg drifting away an idealized ice shelf (see \citet{Stern2017} for more details).

%This iceberg drifting simulation is described in a separate study \citep{Stern2017}.
%Here, we instead focus comparing the Lagrangian model to existing Eularian model in a static ice shelf configuration. 
However, before analyzing the improved capabilities of the Lagrangian ice-shelf model, it is necessary to thoroughly test and benchmark the Lagrangian model in a more familiar configuration. 
To do this we compare the Lagrangian ice-shelf-cavity model to existing Eularian model in a static ice-shelf configuration. 
For this comparison, we use the model configuration developed as part of the Marine Ice Ocean Modeling Inter-comparison Project (MISOMIP). 
The goals of this study are (i) to introduce and describe the Lagrangian ice-shelf model, and (ii) to prove that the Lagrangian model can replicate the behavior of an Eularian ice-shelf model when modeling ocean cavities beneath static ice shelves. 
%In particular, we aim to show that for a static ice shelf, Lagrangian ice-shelf model gives the same results as an Eularian model in the same configuration (using the same ocean model). 
Demonstrating that the Lagrangian model compares well to more traditional Eularian models is a prerequisite for using the Lagrangian model to move beyond the capabilities of the Eularian model, and increases our confidence in more complex Lagrangian simulations involving calving icebergs. 
%An analysis of an ice-shelf cavity with a tabular iceberg drifting away from the from the ice shelf is provided in a separate study \citep{Stern2017}.

@@

\section{Lagrangian model description}
The Kinematic Iceberg Dynamics (KID) model is Lagrangian model that has been developed in order to simulate ice-shelf cavities with breakable ice shelves. The model represents ice shelves using Lagrangian elements joined together by numerical bonds. By breaking these bonds, the model is able to simulate ice-shelf calving and iceberg breakup.

The goal of this study is to show that this Lagrangian ice-shelf model can replicate the results of an Eularian ice-shelf model when used in a static ice-shelf configuration (with all the elements being held stationary). 
Although the ice elements do not move in the static ice-shelf experiment, this configuration is a relevant test for the Lagrangian model, as it tests the interpolation and aggregation between Eularian and Lagrangian grids, and also test the coupling between the ocean and ice-shelf models.
Benchmarking the Lagrangian ice-shelf model to an existing Eularian model in a static ice-shelf configuration is a prerequisite for using the Lagrangian model in more complex experimental setups with moving elements.

In this section we briefly describe the fully dynamic Lagrangian ice-shelf model. We then focus on element packing and interpolation between the Eularian and Lagrangian grids, which are the parts of the model which are relevant to static ice simulations. A complete description of the KID model, including details of the momentum budget and element interactions can be found in \citet{Stern2017}.

\subsection{Kinematic Iceberg Dynamics model}
The KID model is a Lagrangian particle-based model, in that the objects of the model are Lagrangian elements. Each Lagrangian element represents a column of ice that is floating in the ocean. The elements each have their own position, velocity, mass, and a set of dimensions, which can evolve in time.
Each element moves according to its own momentum equation which is solved in the (Lagrangian) reference frame of the element.  The elements are forced by oceanic, sea ice and atmospheric drag forces, as well as a forces due to sea surface height gradients, and the Coriolis force \citep{Bigg1997, Gladstone2001, Martin2010, Stern2017}.  The elements also interact with other elements and can be `bonded' together by numerical bonds, which hold the elements together and allow many elements to move together as a unit. By bonding many ice elements together, the model is able to form larger structures, such as tabular icebergs or ice shelves (see schematic Figures \ref{fig:Schematic}).
% The interactive forces between elements aggregate to give the larger structures an effective rheology.

The Lagrangian elements melt when they are exposed to above-freezing ocean mixed-layer temperatures. The melt rates of the elements are prescribed using parametrizations developed for iceberg and ice-shelf decay \citep{Bigg1997, Holland1999, Gladstone2001, Martin2010}. As the elements melt, heat, salt and mass fluxes are passed to the ocean to simulate freshwater injection to the ocean surface.
% In this study we focus on comparing the model to a Eularian model, so the melt rate of elements is determined using the three equation model for basal melt, which is a typical melting parametrization used beneath ice shelves \citep{Holland1999}.

\subsection{Initializing element geometry and packing}
The elements in the Lagrangian model are assumed to have surface areas which are shaped as equally-sized regular hexagons. We initialize the Lagrangian model by positioning the elements in a staggered lattice, so that the elements fit together and perfectly tile the ice-shelf surface (Figures \ref{fig:Schematic}b). Hexagonal elements are used so that when adjacent pairs of elements are bonded together, the network of bonds form equilateral triangles, which gives rigidity to the larger structure \citep{Stern2017}.

@@

In the experiments presented in this study, the ice-shelf model is coupled to an ocean model, which is runs on a static Eularian grid. At every time step, ice-shelf fields are passed from the ice-shelf model to the ocean model and from the ocean model to the ice-shelf model. Four ocean (mixed layer) fields are passed from the ocean model to the ice-shelf model: temperature, salinity and zonal and meridional velocities. These fields are interpolated onto the Lagrangian grid (i.e.: onto the elements in the ice-shelf model) using a bilinear interpolation scheme.

At the end of a ice-shelf model time step, ice-shelf fields are aggregated from the elements back onto the Eularian ocean grid, and are then passed from the ice-shelf model to the ocean model. The aggregation from the Lagrangian grid to the Eularian ocean grid is done in a way which is consistent with the shape of the elements. For the purposes of aggregation, we assume that the elements have surface areas that are shaped as regular hexagons. Element's properties are aggregated from the Lagrangian grid to the Eularian grid by exactly calculating fraction of an elements surface area that is contained in each ocean grid cell, and dividing the element's properties between the ocean grid cell in proportion to that fraction.

\subsection{Lagrangian vs Eularian ice-shelf model}
In the next section, simulations using the Lagrangian ice-shelf model are compared to simulations using an Eularian ice-shelf model with an an identical configuration. While the internal framework and grids of the Eularian and Lagrangian ice-shelf models are quite different, both models are coupled to the ocean model using the same coupling structure.
Five fields are passed from the ice-shelf models to ocean model: mass, surface area, temperature flux, salinity flux and mass flux. The Eularian model uses the same grid as the ocean model, while in the Lagrangian model these fields are aggregated onto the ocean model grid before being passed through the coupler. Once in the ocean model, these fields and are used to: (i) apply a pressure to the ocean surface, (ii) alter the upper-ocean boundary condition to reflect that the ocean is covered by ice, and (iii) apply salt, temperature and mass fluxes associated with ice-shelf melting and freezing.

Since the coupling with the ocean model is handled in the same way for both the Eularian and Lagrangian models and both models are coupled to the same ocean model, any differences between the two models in the static ice-shelf configuration (using the same melt rate parametrization) are likely due to the interpolation/aggregation between the ice and ocean grids. By accurately aggregating fields from the Lagrangian grid to Eularian ocean grids, we can make the Lagrangian ice-shelf model behave like the Eularian ice-shelf model in static ice-shelf simulations. 
%Apart from the step of aggregating between the Lagrangian and Eularian grids, the interaction between the Lagrangian ice-shelf model and ocean model is exactly the same as the interaction between an Eularian ice-shelf model and an ocean model.

 \section{Experiment Setup}
% the proof of concept for
The KID model is used to simulate the flow within an idealized ice-shelf cavity. The simulations are repeated using an Eularian ice shelf with an identical configuration. The details of the experimental setup for both the Lagrangian and Eularian models are presented here.

\subsection{Domain configuration}
In order for our simulations to be easily comparable to previous models of ice-shelf cavities, we use an experimental setup based on the configuration created for the Marine Ice Ocean Modeling Inter-comparison Project (MISOMIP) \citep{Asay-Davis2016}.
The MISOMIP configuration was developed as a standardized configuration to allow for a comparison between various ocean-ice coupled models. The configuration consists of an idealized ice shelf in a rectangular domain $L_{x}=80$km long and $L_{y}=480$km wide. The ice shelf is grounded on the southern side of the domain with the ice-shelf front at y=650km. The ice thickness and bottom topography of this setup are shown in Figure \ref{fig:ISOMIP_mass_and_topog}. The configuration is the same as that of the Ocean0 setup in the MISOMIP, with three changes made:
\begin{enumerate}
\item The `calving criteria' used in the MISOMIP study (which states that all points in the ice shelf with thickness less than 100m are set to zero thickness) has not been used.
\item The ice shelf has been thickened on the flanks of the domain, so that the latitude of the grounding line increases away from the center of the ice shelf.
\item The ice shelf is configured to be symmetric about its meridional center line ($x=\frac{L_{x}}{2}$). This was achieved by using the average of the left and right flanks of the ice-shelf thickness. 
\end{enumerate}
These three changes were made in order to make the circulation beneath the ice shelf easier to interpret.

\subsection{Ocean Model}
The Lagrangian and Eularian ice shelves are coupled to the MOM6 ocean model \citep{Hallberg2013}. 
The ocean model is run in using vertical coordinate system which is a hybrid between a sigma-level and a z-level coordinate, achieved using ALE regridding-remapping scheme \citep{White2009}. In this vertical coordinate, model layers bend underneath surface topography (i.e.: the ice shelf), as they would in a sigma coordinate model, and intersect the bottom topography, as they would in a z-coordinate model. The model has 72 vertical layers and has a horizontal resolution of $\Delta x= 2$ km. The numerical simulations were all repeated using an isopycnal coordinate (without ALE regrinding-remapping). The results were qualitatively similar to the ALE results, and are therefore not presented here.
% All results presented below are from the ALE simulations, unless specified otherwise.

The ocean parameters used in the simulations are as specified in the MISOMIP configuration \citep{Asay-Davis2016}, and are shown in Table 1. The simulation is initialized from rest, with horizontally uniform initial ocean temperature and salinity profiles which vary linearly between specified surface and bottom values: $T_{top}=-1.9^{\circ}$C, $T_{bottom}=1.0^{\circ}$C, $S_{top}=33.8$ psu, $S_{bottom}=34.7$. The maximum ocean depth is $H_{ocean}=720$ m.  A sponge layer is used on the northern boundary, which relaxes back to the initial temperature and salinity with a relaxation time scale of $T_{sponge}=0.1$ days over a distance of 10 km. Melting is set to zero for ocean cells where the ocean column thickness is less than 10m.

@@

%\subsection{Eularian ice shelf simulation:}
\subsection{Ice-shelf decay}
The melt rates in both the Lagrangian and Eularian ice-shelf simulations are calculated using the 3 equation model for ice-shelf decay \citep{Holland1999}.  In both experiments the ice shelf is held stationary. In the Eularian code, this is achieved by setting the ice-shelf velocity to zero, while in Lagrangian ice shelf, the element velocities are set to zero. In both simulations, the ice shelf is thermodynamically active and is able to `melt' but has a constant thickness (as specified in the Ocean0 experiment in the MISOMIP \citep{Asay-Davis2016}). In this setup, ice-shelf melting generates temperature and salinity fluxes into the ocean, but does not change the thickness of the ice shelf / ice elements. This can be thought to represent an ice shelf in dynamic equilibrium where the melt is exactly balanced ice-shelf advection. 
%The simulation is run for 5 years, and results are considered after the spinup period.

@@

@@

This study presents a new Lagrangian framework for modeling sub-ice-shelf cavities. In this framework, the ice shelf is constructed out of many Lagrangian elements, which are bonded together by numerical bonds. The collection of Lagrangian elements can be thought of as being a Lagrangian grid, which stores the ice-shelf properties. Unlike Eularian grids, the nodes (elements) can move at every time step, altering the shape of the ice shelf. This allows us to use the Lagrangian ice-shelf model to large pieces of the ice-shelf breaking away from the ice shelf to becoming tabular icebergs that are fully embedded in the ocean (Figures  \ref{fig:Iceberg_calving}). This capability is currently not possible using more traditional Eularian models \citep{Stern2017}.
%
%A consequence of using a Lagrangian ice-shelf model is all variables need to be interpolated from the Lagrangian grid onto the Eularian ocean grid in order to be passed to the ocean model. This is done by exactly calculation the fraction of each ice element that lies in each ocean grid cell, and dividing the element's properties between the ocean grid cell in proportion to that fraction.

>>>>

The current generation of ice-ocean models is unable to represent ice-shelf calving in a physically realistic way despite its controlling role of the mass balance and extent of Antarctic ice shelves.
This is primarily due to static Eulerian approach of the current ice-shelf/sub-ice-shelf cavity models that does not allow separation and modification of computational domains that would represent breaking of pieces of an ice shelf that would become freely floating icebergs.
In this study we develop a new ice-shelf cavity model where the ice shelf is constructed out of Lagrangian elements that are bonded together by numerical bonds. This Lagrangian framework allows for large pieces of the ice shelf to break away and become tabular icebergs. 
We test this Lagrangian ice -shelf model by simulating the circulation within a (static) idealized ice-shelf cavity, which was developed as part of the Marine Ice Ocean Modeling Inter-comparison Project (MISOMIP). 
The Lagrangian model results are compared to results of an Eularian model simulations with identical configuration.
Results show that the Lagrangian and Eularian models are almost indistinguishable.
% For example, differences in the models are two orders of magnitude smaller than the differences observed when changing the ocean model coordinate system. 
The similarly between the Eularian and Lagrangian models in a static ice-shelf configuration provides a proof of concept for the Lagrangian model, and means that we can confidently use Lagrangian ice-shelf models to extend the capabilities of ice-shelf-cavity models.% The Lagrangian model is then used to simulate a tabular iceberg calving away from the ice shelf and drifting into the open ocean.

@@

%  and tables.
\begin{article}
%% ------------------------------------------------------------------------ %%
%
%  TEXT
%
%% ------------------------------------------------------------------------ %%
\section{Introduction}
%\sout{Floating ice shelves cover vast regions of the Antarctic polar oceans. These massive platforms of ice extend deep into the water column, depressing the surface of the liquid ocean, often to hundreds of meters below global-mean sea level. Beneath the ice shelves, both the bottom topography and the ice-shelf geometry play a role in steering ocean currents [Nicholls1996, Jenkins2010,  Stern2014]. The topographic constraint imposed by the ice shelves at the ocean's upper boundary significantly affects the circulation within the ice-shelf cavities.}
%
Satellite observations show that ice-shelf decay occurs via two main processes: melting and breaking \citep{Depoorter2013, Rignot2013}. Each of these is responsible for approximately half of the ice-shelf decay, and each influences the surrounding ocean (and ice-shelf geometry) in a distinct way.
Melting at the base of ice shelves causes fluxes freshwater into the ice-shelf cavity. The input of buoyant meltwater creates rising density plumes, which are guided along the ice-shelf base, and help drive ocean circulation beneath the ice shelves \citep{MacAyeal1984,  Holland2006}. Over time, melting at the ice-shelf base can erode the ice shelf, gradually altering the ice-shelf geometry.
In contrast, iceberg calving causes sudden changes to the ice-shelf geometry, and releases giant icebergs into the ocean. After calving, these tabular icebergs can travel large distances and impact ocean hydrography \citep{Martin2010, Stern2015}, sea-ice formation \citep{Robinson2012, Stern2016} and ocean biology \citep{Smith2007, Vernet2012, Biddle2015} many miles away.

%\sout{In recent years society's need for improved projections of future sea level has lead to an increased focus on the Antarctic mass balance, and hence, on modeling Antarctic sub-ice-shelf cavities [Asay-Davis2016]}
Modeling the ocean beneath the ice shelves presents a unique set of challenges, since (i) the presence of ice shelves provides a quasi-rigid upper boundary for the ocean model which is not encountered elsewhere in the ocean, and (ii) melting and breaking ice shelves imply changing ocean boundary conditions which present numerous numerical difficulties.

The earliest models of ocean ice-shelf cavities were developed using static ice shelves with a fixed shape \citep{Hellmer1989, Determan1994, Grosfeld1997, Holland2001, Losch2008}. In these models, ice-shelf melting was represented through salinity and temperature fluxes, while the ice-shelf geometry remained unchanged. Later models of ice-shelf cavities allowed the ice-shelf geometry to evolve as the ice shelf melted, permitting the study of coupled ocean-ice phenomena \citep{Gladish2012, Sergienko2013}. More recently, dynamic ice-shelf models have been coupled to the ocean cavity, allowing the study of grounding line migration which is of key importance for sea level rise projections \citep{Grosfeld2004, Goldberg2012, De_Rydt2006, Seroussi2017}.

All models of ice-shelf cavities to date have omitted ice-shelf breaking and iceberg detachment. This is because (i) there is much uncertainty about the physics that govern ice-shelf breaking \citep{Benn2007, Alley2008, Levermann2012, Bassis2013}, and (ii) current models of ice-shelf cavities represent the ice shelves on static Eularian grids, which do not lend themselves to modeling iceberg detachment and drift. In contrast, existing {\it iceberg} models represent icebergs as Lagrangian particles, since this is a convenient way to model discrete objects traveling over large distances \citep{Bigg1997, Gladstone2001, Martin2010, Marsh2015}. To date there has been no real effort to synthesize these two approaches (i.e.: to combine ice shelf and iceberg models).

In this study we develop a new ice-shelf cavity model where the ice shelf is simulated with Lagrangian elements \citep{Stern2017}. In this model, the ice shelf is constructed out of Lagrangian elements which are bonded together by numerical bonds (see schematic Figures \ref{fig:Schematic}). This Lagrangian framework allows for large pieces of the ice shelf to break away and become tabular icebergs. An example of this enhanced capability of the Lagrangian model is demonstrated in Figures \ref{fig:Iceberg_calving}, which shows a tabular iceberg drifting away from an idealized ice shelf (see \citet{Stern2017} for more details).

%This iceberg drifting simulation is described in a separate study \citep{Stern2017}.
%Here, we instead focus comparing the Lagrangian model to existing Eularian model in a static ice shelf configuration. 
However, before analyzing the improved capabilities of the Lagrangian ice-shelf model, it is necessary to thoroughly test and benchmark the Lagrangian ice-shelf model in a more familiar configuration. 
To do this we compare the Lagrangian ice-shelf-cavity model to existing Eularian model in a static ice-shelf configuration. 
For this comparison, we use the model configuration developed as part of the Marine Ice Ocean Modeling Inter-comparison Project (MISOMIP). 
The goals of this study are (i) to introduce and describe the Lagrangian ice-shelf model, and (ii) to prove that the Lagrangian model can replicate the behavior of an Eularian ice-shelf model when modeling ocean cavities beneath static ice shelves. 
%In particular, we aim to show that for a static ice shelf, Lagrangian ice-shelf model gives the same results as an Eularian model in the same configuration (using the same ocean model). 
Demonstrating that the Lagrangian model compares well to more traditional Eularian models is a prerequisite for using the Lagrangian model to move beyond the capabilities of the Eularian model, and increases our confidence in more complex Lagrangian simulations involving calving icebergs. 
%An analysis of an ice-shelf cavity with a tabular iceberg drifting away from the from the ice shelf is provided in a separate study \citep{Stern2017}.

@@

\section{Lagrangian model description}
The Kinematic Iceberg Dynamics (KID) model is a Lagrangian model that has been developed in order to simulate ice-shelf cavities with breakable ice shelves. The model represents ice shelves using Lagrangian elements joined together by numerical bonds. By breaking these bonds, the model is able to simulate ice-shelf calving and iceberg breakup.

%\sout{The goal of this study is to show that this Lagrangian ice-shelf model can replicate the results of an Eularian ice-shelf model when used in a static ice-shelf configuration (with all the elements being held stationary). }
Although the ice elements do not move in the static ice-shelf experiment presented in this paper, this configuration is a relevant test for the Lagrangian model, as it tests the interpolation and aggregation between Eularian and Lagrangian grids, and also tests the coupling between the ocean and ice-shelf models.

In this section we describe the fully dynamic Lagrangian ice-shelf model. We then focus on element packing and interpolation between the Eularian and Lagrangian grids, which are the parts of the model which are relevant to static ice simulations. A complete description of the KID model, including details of the momentum budget and element interactions can be found in \citet{Stern2017}.

\subsection{Kinematic Iceberg Dynamics model}
The KID model is a Lagrangian particle-based model, in that the objects of the model are Lagrangian elements. Each Lagrangian element represents a column of ice that is floating in the ocean. The elements have their own position, velocity, mass, and a set of dimensions, which can evolve in time.
Each element moves according to its own momentum balance which is computed in the (Lagrangian) reference frame of the element.  The elements experience oceanic, sea ice and atmospheric drag forces, as well as a forces due to sea surface height gradients, and the Coriolis force \citep{Bigg1997, Gladstone2001, Martin2010, Stern2017}.  The elements also interact with other elements and can be `bonded' together by numerical bonds, which allow many elements to move together as a unit. By bonding many ice elements together, the model is able to form larger structures, such as tabular icebergs or ice shelves (Figures \ref{fig:Schematic}).
% The interactive forces between elements aggregate to give the larger structures an effective rheology.

The thickness and extent of the Lagrangian elements change due to melting when they are exposed to above-freezing ocean mixed-layer temperatures. The melt rates of the elements are prescribed using parametrizations developed for iceberg and ice-shelf decay \citep{Bigg1997, Holland1999, Gladstone2001, Martin2010}. The corresponding heat, salt and mass fluxes are passed to the ocean to simulate freshwater injection to the ocean surface.
% In this study we focus on comparing the model to a Eularian model, so the melt rate of elements is determined using the three equation model for basal melt, which is a typical melting parametrization used beneath ice shelves \citep{Holland1999}.

\subsection{Initializing element geometry and packing}
The elements in the Lagrangian model are shaped as equally-sized regular hexagons. We initialize the Lagrangian model by positioning the elements in a staggered lattice of equally-size hexagons, so that the elements fit together and perfectly tile the ice-shelf surface (Figures \ref{fig:Schematic}b). Hexagonal elements are used so that when adjacent pairs of elements are bonded together, the network of bonds form equilateral triangles, which gives rigidity to the larger structure \citep{Stern2017}. \textcolor{red}{??? Maybe add some comments about accuracy of continuity of fields like pressure etc on the discrete elements???}

@@

In the experiments presented in this study, the ice-shelf model is coupled to an ocean model, which is runs on a static Eularian grid. At every time step, ice-shelf fields are passed from the ice-shelf model to the ocean model and from the ocean model to the ice-shelf model. Four ocean (mixed layer) fields are passed from the ocean model to the ice-shelf model: temperature, salinity and zonal and meridional velocities. These fields are interpolated onto the Lagrangian elements using a bilinear interpolation scheme.

At the end of a ice-shelf model time step, ice-shelf fields are aggregated from the elements back onto the Eularian ocean grid, and are then passed from the ice-shelf model to the ocean model. The aggregation from the Lagrangian grid to the Eularian ocean grid is done in a way which is consistent with the shape of the elements. For the purposes of aggregation, we assume that the elements have surface areas that are shaped as regular hexagons. Element's properties are aggregated from the Lagrangian grid to the Eularian grid by exactly calculating the fraction of an element that is intersects each ocean grid cell, and dividing the element's properties between the ocean grid cell in proportion to that fraction. \textcolor{red}{??? This part is a bit hard to follow, it needs more explanation. Don't know whether a schematic could help. ???}

\subsection{Lagrangian vs Eularian ice-shelf model}
In the next section, simulations using the Lagrangian ice-shelf model are compared to simulations using an Eularian ice-shelf model with an an identical configuration. While the internal framework and grids of the Eularian and Lagrangian ice-shelf models are quite different, both models are coupled to the ocean model using the same coupling structure.
Five fields are passed from the ice-shelf models to ocean model: mass, surface area, temperature flux, salinity flux and mass flux. The Eularian model uses the same grid as the ocean model, while in the Lagrangian model these fields are aggregated onto the ocean model grid before being passed through the coupler. Once in the ocean model, these fields and are used to: (i) apply a pressure to the ocean surface, \textcolor{red}{??? This condition is specific to MOM6. Some explanations are required either about MOM6 specifics or how this could be handled in general. What if an ocean would be z-coordinate, what would be an equivalent of this condition???}(ii) alter the upper-ocean boundary condition to reflect that the ocean is covered by ice, \textcolor{red}{another ambiguous statement} and (iii) apply salt, temperature and mass fluxes associated with ice-shelf melting and freezing.

Since the coupling with the ocean model is handled in the same way for both the Eularian and Lagrangian models and both models are coupled to the same ocean model, any differences between the two models in the static ice-shelf configuration (using the same melt rate parametrization) are likely due to the interpolation/aggregation between the ice and ocean grids. By accurately aggregating fields from the Lagrangian grid to Eularian ocean grids, we aim to make the Lagrangian ice-shelf model behave like the Eularian ice-shelf model in static ice-shelf simulations. 
%Apart from the step of aggregating between the Lagrangian and Eularian grids, the interaction between the Lagrangian ice-shelf model and ocean model is exactly the same as the interaction between an Eularian ice-shelf model and an ocean model.

 \section{Experiment Setup}
% the proof of concept for
%\sout{The KID model is used to represent the ice shelf in an idealized ice-shelf cavity MOM6 ocean flow. The simulations are repeated using an Eularian ice shelf with an identical configuration. The details of the experimental setup for both the Lagrangian and Eularian models are presented here.}
%
\subsection{Domain configuration}
In order for our simulations to be easily comparable to previous models of ice-shelf cavities, we use an experimental setup based on the configuration created for the Marine Ice Ocean Modeling Inter-comparison Project (MISOMIP) \citep{Asay-Davis2016}.
The configuration consists of an idealized ice shelf in a rectangular domain $L_{x}=80$km long and $L_{y}=480$km wide. The ice shelf is grounded on the southern side of the domain with the ice-shelf front at y=650km. The ice thickness and bottom topography of this setup are shown in Figure \ref{fig:ISOMIP_mass_and_topog}. The configuration is the same as that of the Ocean0 setup in the MISOMIP, with three changes made:
\begin{enumerate}
\item The `calving criteria' used in the MISOMIP study (which states that all points in the ice shelf with thickness less than 100m are set to zero thickness) has not been used.
\item The ice shelf has been thickened on the flanks of the domain, so that the latitude of the grounding line increases away from the center of the ice shelf.
\item The ice shelf is configured to be symmetric about its meridional center line ($x=\frac{L_{x}}{2}$). This was achieved by using the average of the left and right flanks of the ice-shelf thickness. 
\end{enumerate}
These three changes were made in order to make the circulation beneath the ice shelf easier to interpret.

\subsection{Ocean Model}
The Lagrangian and Eularian ice shelves are coupled to the MOM6 ocean model \citep{Hallberg2013}. 
The ocean model is run using a hybrid vertical coordinate system which blends a sigma-level and a z-level coordinate \citep{Stern2017}, implemented using the ALE method \citep{White2009}. In this vertical coordinate, model layers bend underneath surface topography (i.e.: the ice shelf), as they would in a sigma coordinate model, and intersect the bottom topography, as they would in a z-coordinate model. The model has 72 vertical layers and has a horizontal resolution of $\Delta x= 2$ km. The numerical simulations were all repeated using an isopycnal coordinate (without ALE regrinding-remapping). The results were qualitatively similar to the hybrid-coordinate results, and are therefore not presented here.
% All results presented below are from the ALE simulations, unless specified otherwise.

The ocean parameters used in the simulations are as specified in the MISOMIP configuration \citep{Asay-Davis2016}, and are shown in Table 1. The simulation is initialized from rest, with horizontally uniform initial ocean temperature and salinity profiles which vary linearly between specified open-ocean surface and bottom values: $T_{top}=-1.9^{\circ}$C, $T_{bottom}=1.0^{\circ}$C, $S_{top}=33.8$ psu, $S_{bottom}=34.7$. The maximum ocean depth is $H_{ocean}=720$ m.  A sponge layer is used on the northern boundary, which relaxes back to the initial temperature and salinity with a relaxation time scale of $T_{sponge}=0.1$ days over a distance of 10 km. Melting is set to zero for ocean cells where the ocean column thickness is less than 10m.

@@

%\subsection{Eularian ice shelf simulation:}
\subsection{Sub-ice-shelf melting}
The melt rates in both the Lagrangian and Eularian ice-shelf simulations are calculated using the 3 equation model for ice-shelf decay \citep{Holland1999}.  In both experiments the ice shelf is held stationary. In the Eularian code, this is achieved by setting the ice-shelf velocity to zero, while in Lagrangian ice shelf, the element velocities are set to zero. In both simulations, the ice shelf is thermodynamically active and is able to `melt' but has a time-invariant thickness (as specified in the Ocean0 experiment in the MISOMIP \citep{Asay-Davis2016}). In this setup, ice-shelf melting generates temperature and salinity fluxes into the ocean, but does not change the thickness of the ice shelf / ice elements. This can be thought to represent an ice shelf in dynamic equilibrium where the melt is exactly balanced ice-shelf advection. 
%The simulation is run for 5 years, and results are considered after the spinup period.

@@

\textcolor{red}{??? This section is a bit thin. Perhaps it'd be worth looking at more fields and also discuss to under which circumstances differences between the KID and Eularian models can be larger and start affecting simulations and the results ???}

@@

This study presents a new Lagrangian framework for modeling sub-ice-shelf cavities. In this framework, the ice shelf is constructed out of many Lagrangian elements, which are bonded together by numerical bonds. The collection of Lagrangian elements store the ice-shelf properties. Unlike Eularian grids, the nodes (elements) can move at every time step, altering the shape of the ice shelf. This allows us to use the Lagrangian ice-shelf model to large pieces of the ice-shelf breaking away from the ice shelf to becoming tabular icebergs that are fully embedded in the ocean (Figures  \ref{fig:Iceberg_calving}). This capability is currently not possible using more traditional Eularian models \citep{Stern2017}.
%
%A consequence of using a Lagrangian ice-shelf model is all variables need to be interpolated from the Lagrangian grid onto the Eularian ocean grid in order to be passed to the ocean model. This is done by exactly calculation the fraction of each ice element that lies in each ocean grid cell, and dividing the element's properties between the ocean grid cell in proportion to that fraction.

The framework capabilities are demonstrated by modeling the circulation beneath a (static) idealized ice shelf, which was developed as part of the MISOMIP inter-comparison project. The results from the Lagrangian ice-shelf model fit within our paradigm of understanding for circulation within ice-shelf cavities: buoyant meltwater that enters the ice-shelf cavity drives freshwater plumes at the ice-shelf base, which drives the circulation within the cavity.  The circulation, melt rates and ocean hydrography achieved using the Lagrangian ice shelf compare well with other simulations in the MISOMIP experiments.
